---
title: "Classification with and without precautions"
author: "Reka Solymosi"
date: "21/07/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(haven)
library(ggplot2)
library(tidyr)
library(stringr)
library(forcats)
library(nnet)
library(stargazer)
library(broom)
library(MASS)
library(dplyr)
library(networkD3)
library(RColorBrewer)
library(janitor)
library(gridExtra)
library(Hmisc)

# read in data
data <- read_dta("https://www.dropbox.com/s/l0k2if9axtpu8mu/Covid-19_Police_Study_Longitudinal_W1-W3.dta?dl=1")
#filter out people who didn't answer foc-19 Q and choose weight to use throughout
data <- data %>% mutate(weight = ifelse(is.na(w1w2weight), 1, w1w2weight))

data <- data %>% 
  mutate(foc19_s_w3 = case_when(
          # Unworried
          c_foc==2 ~ "Unworried", # not worried
          # Functional
          c_foc==1 & # worried about catching covid in past 3 weeks
          c_focqol < 7   ~ "Functional Worry",  #  quality of life is not reduced by worry
         #Dysfunctional
          c_foc== 1 & # worried about covid
            c_focqol > 6  ~ "Dysfunctional Worry", #quality of life affected by worry
          TRUE ~ NA_character_),    #everyone else 
    foc19_s_w2 = case_when(
          # Unworried
          b_foc==2 ~ "Unworried", # not worried
          # Functional
          b_foc==1 & # worried about catching covid in past 3 weeks
          b_focqol < 7   ~ "Functional Worry",  #  quality of life is not reduced by worry
         #Dysfunctional
          b_foc== 1 & # worried about covid
            b_focqol > 6  ~ "Dysfunctional Worry", #quality of life affected by worry
          TRUE ~ NA_character_),    #everyone else 
         #wave3
         foc19_c_w3 = case_when(
          # Unworried
          c_foc==2 ~ "Unworried", # not worried
          # Functional
          c_foc==1 & # worried about catching covid in past 3 weeks
            c_prec == 1 & # takes precautions
            c_precs > 6 & # feels safer from precs
            c_precqof < 7 & # quality of life is not reduced by precautions
          c_focqol < 7   ~ "Functional Worry",  #  quality of life is not reduced by worry
         #Dysfunctional
          c_foc== 1 & # worried about covid
            b_focqol > 6  ~ "Dysfunctional Worry 1", #quality of life affected by worry
         # or 
          c_foc == 1 &  # worried about covid
           c_prec == 2  ~ "Dysfunctional Worry 2", # and doesn't take precautions
             # or  
           c_foc== 1 &  # worried about covid
             c_prec == 1 & # takes precautions
             c_precqof > 6  ~ "Dysfunctional Worry 3", # and quality of life affected by precautions
         # or  
           c_foc== 1 &  # worried about covid
             c_prec == 1 & # takes precautions
             c_precs < 7  ~ "Dysfunctional Worry 4", #doesnt feel safer from precs
            TRUE ~ NA_character_),
         foc19_c_w2 = case_when(
            # Unworried
            b_foc==2 ~ "Unworried", # not worried
            # Functional
            b_foc==1 & # worried about catching covid in past 3 weeks
            b_focqol < 7  & # quality of life not reduced by worry 
              b_prec == 1 & # takes precautions
              b_precs >= 6 & # feels safer from precs
              b_precqof < 7 ~ "Functional Worry",  #  quality of life is not reduced by precautions
           #Dysfunctional
            b_foc== 1 & # worried about covid
              b_focqol > 6  ~ "Dysfunctional Worry", # and quality of life affected by worry
           # or 
            b_foc== 1 &  # worried about covid
             b_prec == 2  ~ "Dysfunctional Worry 2", # and doesn't take precautions
             # or  
           b_foc== 1 &  # worried about covid
             b_prec == 1 & # takes precautions
             b_precqof > 6  ~ "Dysfunctional Worry 3", # and quality of life affected by precautions
           # or  
           b_foc== 1 &  # worried about covid
             b_prec == 1 & # takes precautions
             b_precs <= 6  ~ "Dysfunctional Worry 4", #doesnt feel safer from precs
            TRUE ~ NA_character_)    #everyone else 
          ) 

# data$foc19_c_w2 <- factor(data$foc19_c_w2, levels = c("Unworried", "Functional Worry","Dysfunctional Worry" ))
# data$foc19_c_w3 <- factor(data$foc19_c_w3, levels = c("Unworried", "Functional Worry","Dysfunctional Worry" ))
# data$foc19_s_w2 <- factor(data$foc19_s_w2, levels = c("Unworried", "Functional Worry","Dysfunctional Worry" ))
# data$foc19_s_w3 <- factor(data$foc19_s_w3, levels = c("Unworried", "Functional Worry","Dysfunctional Worry" ))

data <- data %>% 
  mutate(race_coded = as_factor(race), 
         race_coded = case_when(
    grepl("Asian or Asian British", race_coded) |
    grepl("Black or Black British", race_coded) |
    grepl("Mixed", race_coded) ~ "BAME",
    grepl("White", race_coded) ~ "White", 
    TRUE ~ "Other or did not say"), 
    gender = as_factor(gender)) 


```

## Complex classification with different routes to dysfunctional fear

- **Dysfunctional 1** group are those who are quality of life affected by worry
- **Dysfunctional 2** group are those who's quality of life is not affected by worry but they don't take precautions
- **Dysfunctional 3** group are those who's quality of life is not affected by worry and they take precautions but these precautions then affect their quality of life 
- **Dysfunctional 4** group are those who's quality of life is not affected by worry and they take precautions but these precautions do not make them feel safer

#### Wave 2

```{r}
data %>% group_by(foc19_c_w2) %>% count()

```

#### Wave 3
```{r}
data %>% group_by(foc19_c_w3) %>% count()

```



If we sum them all then


#### Wave 2

```{r}
data %>% 
  mutate(foc19_c_w2 = ifelse(grepl("Dysfunc", foc19_c_w2), "Dysfunctional Worry", foc19_c_w2)) %>% 
  group_by(foc19_c_w2) %>% count()

```

#### Wave 3

```{r}
data%>% 
  mutate(foc19_c_w3 = ifelse(grepl("Dysfunc", foc19_c_w3), "Dysfunctional Worry", foc19_c_w3)) %>% group_by(foc19_c_w3) %>% count()

```

## Simple classification

In this case we classify based on only whether or not worry affects quality of life, and don't take into account the precautions in the classification:

#### Wave 2

```{r}
data %>% group_by(foc19_s_w2) %>% count()

```

#### Wave 3
```{r}
data %>% group_by(foc19_s_w3) %>% count()

```


## Run the models with the different classifications

### 







